% To familiarize yourself with this template, the body contains
% some examples of its use.  Look them over.  Then you can
% run LaTeX on this file.  After you have LaTeXed this file then
% you can look over the result either by printing it out with
% dvips or using xdvi.
%

\documentclass[twoside]{article}
%\usepackage{soul}
\usepackage{./lecnotes_macros}


\begin{document}
%FILL IN THE RIGHT INFO.
%\lecture{**LECTURE-NUMBER**}{**DATE**}{**LECTURER**}{**SCRIBE**}
\lecture{13}{9 October 2023}{Shashank Vatedka}{Gautam Singh}
%\footnotetext{These notes are partially based on those of Nigel Mansell.}

%All figures are to be placed in a separate folder named ``images''

% **** YOUR NOTES GO HERE:

\section{Encoding Sparse Vectors}

Suppose that \(\vec{x}\in\bR^d\) is sparse and has \(s\) nonzero entries. To
encode \(\vec{x}\), we send the locations of the nonzero coordinates using
\(\log_2\binom{d}{s}\) bits and separately encode the nonzero values.

The \emph{Elias code} is used to encode a positive integer \(K\). In this scheme
of codes, we need \(\ceil{\log_2 K}\) bits for the integer, and prepend it with
\(\ceil{\log_2 \ceil{\log_2 K}}\) bits for the length, and so on. The first set
will be a \emph{unary code}, where any positive integer \(L\) is encoded as
\(0^{L}1\). The final 1 is to delimit the unary code.

The Elias code is faster than entropy-based codes such as Huffman or
Arithmetic codes. Thus, it is preferred for cases like SGD where the time
complexity is important.

\section{Shared Randomness Among Users}

Suppose that two users \(U_1\) and \(U_2\) use randomized rounding, where 
\(U_2 = 1 - U_1\). Then, \(U_1\) and \(U_2\) are identically distributed.

Suppose that \(x_1 = x_2 = x\). The MSE is then

\begin{align}
    \mean{\brak{\frac{\hat{x_1} + \hat{x_2}}{2}-x}^2} &= \frac{1}{4}\mean{\brak{\brak{\hat{x_1}-x}+\brak{\hat{x_2}-x}}^2} \\
                                                      &= \frac{1}{4}\brak{2x\brak{1-x}+2\mean{\brak{\hat{x_1}-x}\brak{\hat{x_2}-x}}} \\
                                                      &= \frac{1}{2}\brak{x\brak{1-x}+\sbrak{2x-1}_+ - x^2}
                                                      \label{eq:mse-shared-rand}
\end{align}

where we have

\begin{align}
    \mean{\hat{x_1}\hat{x_2}} &= \mean{\mathbbm{1}_{\cbrak{U_1 < x_1}}\mathbbm{1}_{\cbrak{U_2 < x_2}}} \\
                              &= \pr{U_1 < x_1,\ U_2 < x_2} \\
                              &= \pr{1 - x < U_1 < x} = \sbrak{2x-1}_{+}
                              \label{eq:mean-prod}
\end{align}

Observe that this MSE is better than that produced by two iid uniform random
variables.

\end{document}
